---
title: "QC Report - NONMEM PK Analysis Dataset"
subtitle: "CDISCPILOT01 (Xanomeline)"
author: "H. Besbassi"
date: ""
output:
  html_document:
    toc: true
    toc_float: true
    theme: flatly
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, fig.width = 10, fig.height = 6)
library(tidyverse)
library(knitr)
library(DT)
```

# 1. Dataset Overview

```{r load}
nm_data <- read_csv("../outputs/nm_dataset.csv", show_col_types = FALSE, na = ".")
qc_results <- read_csv("../outputs/qc_checklist.csv", show_col_types = FALSE)

obs_data <- nm_data %>% filter(EVID == 0)
dose_data <- nm_data %>% filter(EVID == 1)
subj_data <- nm_data %>% distinct(ID, .keep_all = TRUE)
```

| Metric | Value |
|--------|-------|
| Total records | `r nrow(nm_data)` |
| Subjects | `r n_distinct(nm_data$ID)` |
| Variables | `r ncol(nm_data)` |
| Dosing records | `r nrow(dose_data)` |
| Observations | `r nrow(obs_data)` |
| BLQ observations | `r sum(obs_data$BLQ == 1)` (`r round(100*sum(obs_data$BLQ==1)/nrow(obs_data), 1)`%) |
| Time range | `r min(nm_data$TIME, na.rm=TRUE)` to `r max(nm_data$TIME, na.rm=TRUE)` hours |

# 2. QC Checklist Results

```{r qc_table}
qc_results %>%
  mutate(status = case_when(
    status == "PASS" ~ "PASS",
    status == "FAIL" ~ "**FAIL**",
    status == "WARNING" ~ "WARNING",
    status == "INFO" ~ "INFO",
    TRUE ~ status
  )) %>%
  kable()
```

**Overall QC Status: `r if_else(sum(qc_results$status == 'FAIL') == 0, 'QC PASS', 'QC FAIL')`**

# 3. Missingness Analysis

```{r missingness}
miss_summary <- nm_data %>%
  summarise(across(everything(), ~sum(is.na(.)))) %>%
  pivot_longer(everything(), names_to = "variable", values_to = "n_missing") %>%
  mutate(pct_missing = round(100 * n_missing / nrow(nm_data), 1)) %>%
  filter(n_missing > 0) %>%
  arrange(desc(n_missing))

if (nrow(miss_summary) > 0) {
  ggplot(miss_summary, aes(x = reorder(variable, n_missing), y = pct_missing)) +
    geom_col(fill = "#2E75B6") +
    coord_flip() +
    labs(title = "Missing Values by Variable", x = "", y = "% Missing") +
    theme_minimal()
} else {
  cat("No missing values in dataset.")
}
```

# 4. BLQ Analysis

```{r blq}
blq_time <- obs_data %>%
  mutate(time_bin = cut(TAD, breaks = c(-Inf, 0, 2, 6, 12, 24, 48, Inf),
                        labels = c("Pre-dose", "0-2h", "2-6h", "6-12h", "12-24h", "24-48h", ">48h"))) %>%
  group_by(time_bin) %>%
  summarise(pct_blq = round(100 * mean(BLQ == 1, na.rm = TRUE), 1),
            n = n(), .groups = "drop")

ggplot(blq_time, aes(x = time_bin, y = pct_blq)) +
  geom_col(fill = "#95A5A6") +
  geom_text(aes(label = paste0("n=", n)), vjust = -0.5, size = 3) +
  labs(title = "% BLQ by Time After Dose", x = "", y = "% BLQ") +
  theme_minimal()
```

# 5. Covariate Distributions

```{r covariates, fig.height=8}
cov_vars <- c("AGE", "WT")
available_covs <- intersect(cov_vars, names(subj_data))

if (length(available_covs) > 0) {
  cov_long <- subj_data %>%
    select(ID, ARM, all_of(available_covs)) %>%
    pivot_longer(all_of(available_covs), names_to = "covariate", values_to = "value") %>%
    filter(!is.na(value))

  ggplot(cov_long, aes(x = ARM, y = value, fill = ARM)) +
    geom_boxplot(alpha = 0.7) +
    facet_wrap(~covariate, scales = "free_y") +
    labs(title = "Baseline Covariates by Treatment Arm", x = "", y = "") +
    theme_minimal() +
    theme(legend.position = "none")
}
```

# 6. PK Concentration Profiles

```{r pk_profiles}
pk_plot <- obs_data %>%
  filter(BLQ == 0, TAD >= 0, TAD <= 50)

ggplot(pk_plot, aes(x = TAD, y = DV, group = factor(ID), color = ARM)) +
  geom_line(alpha = 0.3, linewidth = 0.3) +
  geom_point(alpha = 0.3, size = 0.8) +
  scale_y_log10() +
  labs(title = "Individual PK Profiles (Quantifiable, TAD 0-50h)",
       x = "Time After Dose (hours)", y = "Concentration (ug/mL, log)") +
  theme_minimal() +
  theme(legend.position = "bottom")
```

# 7. Cross-Check vs Source

```{r crosscheck}
dm_source <- read_csv("../data/sdtm/dm.csv", show_col_types = FALSE)
pc_source <- read_csv("../data/sdtm/pc.csv", show_col_types = FALSE)
ex_source <- read_csv("../data/sdtm/ex.csv", show_col_types = FALSE)

tibble(
  Source = c("DM (subjects)", "PC (PK records)", "EX (dose records)"),
  `Source Count` = c(n_distinct(dm_source$USUBJID), nrow(pc_source), nrow(ex_source)),
  `Dataset Count` = c(n_distinct(nm_data$ID), nrow(obs_data), nrow(dose_data)),
  Notes = c("Excluded 86 placebo/screen failure subjects",
            "Excluded 1016 urine + 1204 non-active-arm observations",
            "Excluded 226 placebo zero-dose records")
) %>% kable()
```

# 8. QC Sign-Off

| Field | Value |
|-------|-------|
| QC Analyst | H. Besbassi |
| QC Date | - |
| Dataset | nm_dataset.csv |
| Overall Status | `r if_else(sum(qc_results$status == 'FAIL') == 0, 'QC PASS', 'QC FAIL')` |
| Checks Run | `r nrow(qc_results)` |
| Failures | `r sum(qc_results$status == 'FAIL')` |
